---
title: "Exercise_RBasics"
author: "Amy"
date: "082625"
output: html_document
---

# Part 1:

```{r}
var1 = 9
var2 = 6
```


 1. Arithmetic Operators

 A. Find the sum of 8 and 8
```{r}
8 + 8
```

 B. Find the product of 2 and 7
```{r}
2 * 7
```

 C. Exponentiate 3 to the power of 3
```{r}
3^3
```

 D. Find the difference of 89 and 16
```{r}
89 - 16
```

 E. Find the quotient of 9 and 2

```{r}
9 %% 2
```

 2. Comparison Operators

 A. Write a test to evaluate whether `var1` is less than `var2`
```{r}
var1 < var2
```

 B. Write a test to evaluate whether `var1` is equal to `var2`
```{r}
var1 == var2
```

 C. Write a test to evaluate whether the sum of `var1` and `var2` is greater than or equal to the product of `var1` and `var2`
```{r}
(var1 + var2) >= (var1 * var2)
```

 D. Write a test to evaluate whether `var2` squared is not equal to `var1` cubed.
```{r}
(var2^2) != (var1^3)
```

 E. Write a test to evaluate whether the remained of `var1` squared divided by `var2` is greater than 0.

```{r}
((var1^2) %% var2) > 0
```

 F. Write a test to evaluate whether `var2` squared is less than or equal to 50

```{r}
(var2^2) <= 50
```

 3. Logical Operators

 A. Use `&` to evaluate 2A. and 2C.
```{r}
(var1 < var2) & ((var1 + var2) >= (var1 * var2))
```

 B. Use `!` to evaluate 2D.
```{r}
!((var2^2) != (var1^3))
```

 C. Use `|` to evaluate 2E. and 2B.
```{r}
(((var1^2) %% var2) > 0) | (var1 == var2)
```

 D. Now, use each operator (`&`, `!`, and `|`) combined with at least one other operator (logical or comparison operators).

`&`

```{r}
(var1 * 2) & (36 - var2)
```

`!`

```{r}
!(var1 == var2)
```

`|`

```{r}
(var1 == 9) | (var2 > 10)
```

 4. Define one variable that contains the number 3, another containing "Hi!", and another containing `FALSE`. Use `typeof()` to verify the type of each of these variables.

```{r exercise 4}
var_num <- 3
var_char <- "Hi!"
var_log <- FALSE

typeof(var_num)
typeof(var_char)
typeof(var_log)
```


 5. Create 3 vectors that each contain a different type of data (number, character, and logical). See which operations can be applied to each vector.

```{r exercise 5}
vector1 <- c(1.1, 3.4, 5.6, 7.8)
vector1

vector1 * 9
vector1 <= 6

vector2 <- c("Matcha", "Boba", "Coffee", "Smoothie")
vector2

vector2 != "Boba"

vector3 <- c(TRUE, FALSE, FALSE, TRUE)
vector3

vector3 == FALSE
vector3 * 2
```

 6. Index and save the first element from your first vector to a variable called `first_first`, and the last element from your last vector to one called `last_last`.

```{r exercise 6}
first_first <- vector1[1]

last_last <- vector3[4]

first_first
last_last
```

 7. Complete the code below to create a dataframe called `my_df` that has information about you.

```{r exercise 7}
 my_df <- data.frame("Name" = "Amy",
               "Course" = "RBootcamp",
               "Term" = "Summer25")
my_df
```

 8. 
 
 A. Index the first row of your dataframe and save it to a variable called `df_first_row`. Index the course name (1st value of 2nd row) and save it to a variable called `course_name`. 
 
```{r exercise 8a}
df_first_row <- my_df[1,]
course_name <- my_df[1, 2]

df_first_row
course_name
```
 
 B. Now index the course name in a different way and save it to a variable called `course_name_alt`. Write a test to valuate whether `course_name` and `course_name_alt` are equivalent.

```{r exercise 8b}
course_name_alt <- my_df$Course
course_name_alt

course_name == course_name_alt
```


# Part 2:

 1. Install the packages listed below:

- tidyverse
- patchwork
- magrittr
- viridis
- RColorBrewer
- DT
- ggExtra
- gridExtra
- reprex
- knitr
- markdown
- palmerpenguins
- janitor
- ggridges

```{r, eval = FALSE}

install.packages("tidyverse")
install.packages("patchwork")
install.packages("magrittr")
install.packages("viridis")
install.packages("RColorBrewer")
install.packages("DT")
install.packages("ggExtra")
install.packages("gridExtra")
install.packages("reprex")
install.packages("markdown")
install.packages("palmerpenguins")
install.packages("janitor")
install.packages("ggridges")

library("tidyverse")
library("patchwork")
library("magrittr")
library("viridis")
library("RColorBrewer")
library("DT")
library("ggExtra")
library("gridExtra")
library("markdown")
library("palmerpenguins")
library("janitor")
library("ggridges")

```


 2. Create a new code chunk below. Name the code chunk "exercise 2".
```{r exercise 2}




```

 3. Clean up all the code below so that it follows the proper formatting guidelines listed above.

```{r exercise 3, eval = FALSE}
df[, 4]
df[, 2]
df[, 0]

mean(df$col, na.rm = TRUE)
mean(df$col, na.rm = TRUE)

3 + 4 + 5
-2 * (8^2)

#compute persons height
height <- feet * 12 + inches
(T + 1) < 2
!(F | F + 1)

long_computation(
  'little',
  bunny = 1,
  foo * 2,
  hopping,
  through = 'forest')

```

 4. Rewrite the following code as properly formatted pipe chains.

```{r exercise 4, eval = FALSE}

pull(swoop(loop(laces)))

loop(laces) %>%
  swoop() %>%
  pull(laces)

leave_house(get_dressed(get_out_of_bed(wake_up(me))))

me %>%
  wake_up() %>%
  get_out_of_bed() %>%
  get_dressed() %>%
  leave_house()
  
```

 5. Download the files for today's class that came with the assignment. Import both the .csv and .RData files. Make sure you save them to variables! Copy/paste their file path in the cell below.

```{r exercise 6, eval = FALSE}
install.packages("readr")
library(readr)

read_csv("C:/Users/amyng/Downloads/dataset.csv")

load("C:/Users/amyng/OneDrive/Documents/GitHub/codingNCML_Wocs/Amy/R Bootcamp/R Basics - 1/R Basics/data-2/data/dataset.RData")
```

 6. Export your `my_df` dataframe as a .csv file called "my_dataframe.csv". Export your dataframe and all the variables you saved to an .RData file called "my_data.RData".

```{r exercise 7, eval = FALSE}
write_csv(my_df, file = "my_dataframe.csv")

save(my_df, file = "my_data.RData")

```


# Feedback 

Please save the appropriate number to the variables below:

1. Approximately how long (in minutes) do you think it took you to complete this assignment?

```{r}
timeEstimate = 120
```

2. How difficult or challenging did you find this assignment?

Extremely easy -------------------------------------- Extremely challenging 
                                                     (impossible to complete)
    1                                                           10
    
```{r}
challenge = 5
```
    
    
3. How engaging or fun did you find this assignment?
  
Extremely Boring and Dry -------------------------------------- Extremely Engaging and Fun 
          1                                                                   10

```{r}
engagement = 8
```

4. If you have any other feedback about this assignment, please provide it below.




