---
title: "RBasics_Practice"
author: "Amy"
date: "082625"
output: html_document
---


1. Use each of the following operators in a new expression:

"+"

```{r}
5 + 4
```

"-"

```{r}
10 - 7
```

"*"

```{r}
7 * 7
```

"/"

```{r}
8 / 2
```

"^"

```{r}
6^2
```

"%%"

```{r}
10 %% 3
```


2. Use each of the following operators in a new expression:

"<"

```{r}
50 < 100
```

">"

```{r}
100 > 50
```

"<="

```{r}
3.6 <= 3.6
```

">="

```{r}
4.5 >= 3.5
```

"=="

```{r}
1 == 1
```

"!="

```{r}
4 != 2
```

3. 
  A. Use each of the operators below in a new expression. 

"&"

```{r}
TRUE & FALSE
```

"|"

```{r}
TRUE | FALSE
```

"!"

```{r}
!TRUE
```

B. Use *each* operator (`&`, `!`, and `|`) combined with at least one operator from exercise 1 or 2 as well.

```{r}
x <- 5
y <- 10

(x + 2 > 6) & (y - 8 == 2)
!(x < y - 3)
(x * 2 > 15) | (y * 5 == 50)
```


C. Now, use *each* operator (`&`, `!`, and `|`) combined with at least one *other* operator (logical or comparison operators).

```{r}
x <- 7
y <- 12

(x > 5) & (y <= 15)
!(x == 7)
(x < 5) | (y >= 8)
```


4. Define variables of each of the following types and use `typeof()` to verify their type.

A. number

```{r}
a <- 6.5
typeof(a)
```

B. character

```{r}
b <- "Boba"
typeof(b)
```


C. logical

```{r}
c <- TRUE
typeof(c)
```

5. Create 3 vectors that each contain a different type of data (number, character, and logical). See which operations can be applied to each vector.

  A. number

```{r}
myVector <- c(1, 2, 3, 4)
myVector

myVector * 6
myVector >= 2
```

  B. character

```{r}
myVector2 <- c("Hi", "Hello", "Hey", "Hola")
myVector2

myVector2 != "Hi"
```

  C. logical

```{r exercise 5}
myVector3 <- c(TRUE, FALSE, FALSE, TRUE)
myVector3

myVector3 + 1
myVector3 == TRUE
```

6.

  A. Index the 1st and 2nd element from the vector in 5A

```{r}
myVector[1]
myVector[2]
```

  B. Index the 2nd and 3rd element from the vector in 5B.
  
```{r}
myVector2[2]
myVector2[3]
```
  
  C. Save the first element from your first vector to a variable called `first_first`.
  
```{r}
first_first <- myVector[1]
first_first
```

  D. Save the last element from your last vector to one called `last_last`.
  
```{r}
last_last <- myVector3[4]
last_last
```
  

7. Create a dataframe called `my_df` that has information about your top 3 favorite animals. It should contain 3 columns: 1. One that has the words, "first", "second", and "third". 2. Another that has the animals names. 3. A third having a rating, on a scale of 1-10, of how likely you think it is you could beat that animal in a fight to the death.

```{r exercise 7}
my_df = data.frame("Rank" = c("First", "Second", "Third"),
                   "Animals" = c("Polar Bear", "Red Panda", "Capybara"),
                   "Battle_Rating" = c("1", "4", "3"))
my_df
```

8. Index the second row of your dataframe and save it to a variable called `df_second_row`. Index the name of your second favorite animal and save it to a variable called `second_fave`. Index the battle likelihood rating for your third favorite animal and save it to a variable called `third_rating`.

```{r exercise 8}
df_second_row <- my_df[2,]
second_fave <- my_df[2,2]
third_rating <- my_df[3,3]

df_second_row
second_fave
third_rating
```

9. Rewrite the following code as properly formatted pipe chains.

```{r, eval = FALSE}
park(drive(start_car(find("keys")), to = "campus"))

"keys" %>%
  find() %>%
  start_car() %>%
  drive(to = "campus") %>%
  park()

brush(tooth_paste(running_water(tooth_brush)), teeth)

tooth_brush %>%
  running_water() %>%
  tooth_paste() %>%
  brush(teeth)
```


# Feedback 

Please save the appropriate number to the variables below:

1. Approximately how long (in minutes) do you think it took you to complete this assignment?

```{r}
timeEstimate = 90
```

2. How difficult or challenging did you find this assignment?

Extremely easy -------------------------------------- Extremely challenging 
                                                     (impossible to complete)
    1                                                           10
    
```{r}
challenge = 4
```
    
    
3. How engaging or fun did you find this assignment?
  
Extremely Boring and Dry -------------------------------------- Extremely Engaging and Fun 
          1                                                                   10

```{r}
engagement = 8
```

4. If you have any other feedback about this assignment, please provide it below.




