---
title: ""
author: ""
date: ""
output: html_document
---

```{r}
library(tidyverse)
library(palmerpenguins)
```


1. Create your own graph using the `penguins` data and pick 2 of the following to put on your x and y axes: `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, or `body_mass_g` (Note: This should *NOT* be code directly from the lecture or notes).

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g))
```


2. Update your graph above by telling ggplot you want it to render a point for each observation in the penguins dataset. Describe the pattern/relationship you see under the code chunk.

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g)) +
  geom_point()
```

As bill length increases in mm, the penguin's body mass in grams also increases resulting in a positive correlation.


3. Think about what patterns you notice in the visualization you created, and test whether `sex` or `species` might explain those patterns. Write a prediction before the code chunk, test that prediction, and then describe what you see.

*If there are two seperate chunks of colors with mapping by sex or species, this means that these factors could explain those patterns.*

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g, 
                       color = species)) +
  geom_point()

penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g, 
                       color = sex)) +
  geom_point()
```

*There are three distinct clusters in the species visualization while the sex visualization is very scattered. This means that the pattern between bill length and body mass can be better explained by species rather than sex.*

4. Add in the line of best fit with `geom_smooth()`, making sure that your code produces a single line that is fit to ***all*** your data.

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g, 
                       color = species)) +
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  geom_point(aes(color = species))
```

5. Use R code to create or recreate the graphs below

A. Make a visualization to investigate the relationship between bill length (on the y-axis) and bill depth (on the x-axis)

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = bill_depth_mm)) +
  geom_point()
```

B. https://drive.google.com/uc?export=view&id=11Ig5Xq3z8CZwn0ElYmw4GNX0MK09Ja5E

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = bill_depth_mm, 
                       color = species)) +
  geom_point()

```

C. Make a visualization to investigate the relationship between body mass and flipper length. In doing so, make the geometric figure rendering the observations appear in blue. 
```{r}
penguins %>%
  ggplot(mapping = aes(y = body_mass_g,
                       x = flipper_length_mm)) +
  geom_point(color = "blue")
```

D. https://drive.google.com/uc?export=view&id=1hcgTg902GsvRNS2SnZxiv-rtFXRpCiAR

```{r}
penguins %>%
  ggplot(mapping = aes(y = body_mass_g,
                       x = flipper_length_mm)) +
  geom_smooth(method = "lm", se = FALSE, color = "black") + 
  geom_point(color = "blue")
```

E. https://drive.google.com/uc?export=view&id=1RIuM6lRIoMdriUrtBySONizo22L9R_Xm

```{r}
penguins %>%
  ggplot(mapping = aes(y = body_mass_g,
                       x = flipper_length_mm, 
                       color = species)) + 
  geom_point(aes(color = flipper_length_mm > 210)) + 
  geom_smooth(method = "lm", se = FALSE, color = "black")
```

6. 
Part 1: Take the code you produced for Exercise 4 and... 

1. Change the shape to be solid color squares with color mapped to the `species` variable. 

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g, 
                       color = species)) +
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  geom_point(aes(color = species), shape = 15)

```

2. Change the shape to be hollow squares with color mapped to the `species` variable. 

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g, 
                       color = species)) +
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  geom_point(aes(color = species), shape = 0)
```

3. Change the shape to be filled squares with color mapped to the `species` variable and fill set with a color of your choosing.

```{r}
penguins %>%
  ggplot(mapping = aes(y = body_mass_g,
                       x = bill_depth_mm)) +
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  geom_point(aes(color = species),
             shape = 22,
             fill = "white")
```

Part 2: Use R code to recreate the graphs below

A. https://drive.google.com/uc?export=view&id=18OVNxbNI2MhaNTVO7W5FN_Yn-2hkxKgG

```{r}
penguins %>%
  ggplot(aes(y = flipper_length_mm,
             x = bill_length_mm)) + 
  geom_point(aes(color = species,
                 shape = island)) + 
  geom_smooth(method = "lm", se = FALSE, color = "blue")
```

B. https://drive.google.com/uc?export=view&id=1giVWU7yzuD5XND7LE_Ub88tKohtswbtk

```{r}
penguins %>%
  ggplot(aes(y = body_mass_g,
             x = bill_depth_mm,
             color = species)) +
  geom_point(shape = 4)

```

C. https://drive.google.com/uc?export=view&id=1QpR3ph7oCJiLJXRE_hio_agWOQwgNWNd

```{r}
penguins %>%
  ggplot(aes(y = bill_length_mm,
             x = body_mass_g)) + 
  geom_point(color = "red",
             shape = 20) + 
  geom_point(color = "green",
             shape = 25)
```

*Bonus Challenge (ungraded):* See if you can think of a way to use what we demonstrated above to take your code from the Exercise 6, part 2, question 3, and give each observation a thicker outside border. The result should look like the graph below:

https://drive.google.com/uc?export=view&id=1N_9zRcePm8j1ffGGoZgJzUP6XbqisHIR

```{r}
penguins %>%
  ggplot(aes(y = bill_length_mm,
             x = body_mass_g)) + 
  geom_point(color = "green",
             shape = 18,
             size = 6) +
  geom_point(color = "red",
             size = 3,
             shape = 18) 
```

7. Update the main graph you have been working with throughout the exercises to make the observations more translucent. First try doing so by mapping it to `island`. Then, create another graph where you set the alpha yourself. Try to pick the value you think makes the graph most legible and effective.

```{r}
penguins %>%
  ggplot(mapping = aes(y = bill_length_mm,
                       x = body_mass_g, 
                       color = island)) +
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  geom_point(aes(alpha = 1))
```

Feedback 

Please save the appropriate number to the variables below:

1. Approximately how long (in minutes) do you think it took you to complete this assignment?

```{r}
timeEstimate = 60
```

2. How difficult or challenging did you find this assignment?

Extremely easy -------------------------------------- Extremely challenging 
                                                     (impossible to complete)
    1                                                           10
    
```{r}
challenge = 5
```
    
    
3. How engaging or fun did you find this assignment?
  
Extremely Boring and Dry -------------------------------------- Extremely Engaging and Fun 
          1                                                                   10

```{r}
engagement = 8
```

4. If you have any other feedback about this assignment, please provide it below.




