---
title: "Geoms Practice Problem Set"
author: "Amy"
date: "9/10/25"
output: html_document
---

```{r}
library(tidyverse)
# Modify the file path in the code below to properly load the "diamonds_sample.RData" file
load("C:/Users/amyng/OneDrive/Documents/GitHub/codingNCML_Wocs/Amy/R Bootcamp/ggplot expanded - 2/Geoms and Different Graphs/diamonds_sample.RData") 
```

# Intro

You are going to work with a subset of data in the df `diamonds_sample`. All questions below will be in reference to `diamonds_sample`.

* `price` - price in USD
* `carat` - weight of the diamond
* `cut` - quality of the cut
* `color` - diamon color
* `clarity` - a measurement of how clear the diamond is
* `x` - length in mm
* `y` - width in mm
* `z` - depth in mm
* `depth` - total depth percentage
* `table` - width of top of diamond relative to widest point

Below, you are going to practice making several different kinds of graphs, thinking about what type of questions that graph would allow you to answer, what your prediction is, and then summarize what you actually observe. These are to help you with both creating the graphs with R code, but also graph comprehension in general. The latter is a critical skill, without which you will not know what the best graph to make is, even if you have the R skills to make several!

There will be some questions marked below where before running the code to create your graph, you will...
A. Write down *question* that this graph is exploring.
B. Write a brief sentence making a *prediction* of what you expect the relationship or pattern to be. 

Then, after observing the visualization, you will... 
A. Write a brief *summary* describing what that relationship or pattern actually is, and explain what evidence you have for this (e.g., "all the brown colored dots that correspond to 'Ideal' cut diamonds are higher on the y-axis, which here is 'price'. So I can see that 'Ideal' cut diamonds tend to be the most expensive.")

These are clearly marked in the questions that expect them. When a question has a spot that says something like "*Question here*", please delete that text and replace it with your answer.

*NOTE: For some of these visualizations, it may be hard to discern any actual patterns or relationships. Just try your best to describe what you see. If it is too hard to determine anything, state that and explain why!*


1. Use `geom_bar()` to create a bar graph for `clarity`.

```{r}
diamonds_sample %>%
  ggplot(aes(x = clarity)) +
  geom_bar(color = "tomato", fill = "purple")
```

2. Use `geom_bar()` to create a bar graph for `color`. Make the bars themselves be filled with violet and have a blue outline. 

```{r}
diamonds_sample %>%
  ggplot(aes(x = color)) +
  geom_bar(fill = "violet", color = "blue")
```

3. Use `geom_histogram()` to create a histogram for `price`.

```{r}
diamonds_sample %>%
  ggplot(aes(x = price)) +
  geom_histogram(color = "black")
```

4. Use `geom_histogram()` to create a histogram for `carat`. Make the bars themselves be filled with cyan with a turquoise outline. Add a vertical bar at where you think the average might be.

```{r}
diamonds_sample %>%
  ggplot(aes(x = carat)) +
  geom_histogram(fill = "cyan", 
                 color = "turquoise") +
  geom_vline(aes(xintercept = 0.8))

```

5. Use `geom_point()` to create a scatter plot for `carat` and `depth`. Add the line of best fit, and make any aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

Question:
  *What is the relationship between the depth of a diamond and carat?*

Prediction:
  *I predict that higher depth correlates to higher carat, thus, there should be a positive linear line if that is the case. On the other hand, if there is no correlation, there will be flat line. If there is a negative correlation, there will be a negative line of best fit.*

```{r}
diamonds_sample %>%
  ggplot(aes(x = depth, y = carat)) +
  geom_point(color = "blue",
             size = 3,
             alpha = 0.5,
             shape = 15) +
  geom_smooth(method = "lm", se = FALSE)
```

Summary:
  *Based on the graph, there is somewhat of a positive linear relationship between depth and carat. However, it is quite flat, which indiciates that it is not a strong corrrelation.*


6. Create 2D histogram (regular or hexbin) for `y` and `table`.

```{r}
diamonds_sample %>%
  ggplot(aes(x = table, y = y)) +
  geom_hex()
```
Summary:
  *This hexbin indicates that there is greatest amount of diamonds around y of 3.5-8 and table of 55-70. Beyond this chunk, they are sparse.*


7. Take your visualizations from questions 5 and 6 and add information about the individual distributions to the margins (whichever way you prefer).

```{r}
diamonds_sample %>%
  ggplot(aes(x = depth, y = carat)) +
  geom_point(color = "blue",
             size = 2,
             alpha = 0.5,
             shape = 15) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_rug(alpha = .5, size = 2, color = "blue")

library(ggExtra)
p <- ggplot(diamonds_sample, aes(x = table, y = y)) +
  geom_hex() +
  geom_point(alpha = 0)

ggExtra::ggMarginal(p, type = "histogram")
```

Do you think one of these two is better/more effective than the other? If so, which and why? If not, why not?

*I think using geom_rug is better because it allows for an easier way to visualize the data. The other graph is hard to interpret because we cannot see the individual points, and the histograms on the sides make it a bit confusing. *


8. Use `stat_summary()` to show the mean `depth` for `color`. Represent the mean with a bar (note: *not* a vline), and include standard error *bars*. Make any aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

Question:
  *What are the means of a diamonds depth and color?*

Prediction:
  *If depth and color have an influence on each other, the means should vary. If not, the means should all be similar heights.*

```{r}
diamonds_sample %>%
  ggplot(aes(x = color, 
             y = depth)) +
  stat_summary(fun = "mean", geom = "bar") +
  stat_summary(aes(color = color), 
               fun.data = "mean_se", 
               geom = "errorbar", width = 0.5)
```

Summary:
  *The mean values appear to all be the same height. The error bar are very small across the different color of diamonds, showing that depth and color of a diamond are independent of one another.*


9. Use `stat_summary()` to show the mean `price` for `cut`. Represent the mean with a point, and include standard error *lines*. Make any aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = cut, y = price)) +
  stat_summary(aes(color = cut), fun.data = "mean_se", geom = "errorbar") +
  stat_summary(fun = "mean", geom = "point")
```

Summary:
  *Based on the graph, we see that premium cut has the highest price, while ideal is the lowest. Far, good, and very good are generally around the same range.*


10. Take your code from question 9 and modify it so that there is a longdash line connecting each of the points.

```{r}
diamonds_sample %>%
  ggplot(aes(x = cut, y = price)) +
  stat_summary(aes(color = cut), fun.data = "mean_se", geom = "errorbar") +
  stat_summary(fun = "mean", geom = "line", group = 1, linetype = "longdash") +
  stat_summary(fun = "mean", geom = "point")

```

11. `geom_boxplot()` to create a boxplot for `depth` and `clarity`. Make the boxplots sienna with a salmon outline. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = clarity, y = depth)) +
  geom_boxplot(color = "salmon", fill = "grey")
```

What more can you say from this visualization than the one you made in question 8?

*From this visualizion compared to the one in question 8, I can see what the outliers are, the percentiles, and the median. Overall, there is a clearer reading of the individual data points.*


12. `geom_boxplot()` to create a boxplot for `price` and `cut`. Make the boxplots' color and outline (the *fill* and *color* arguments) depend on their `cut`. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph (make sure the median line is visible).

```{r}
diamonds_sample %>%
  ggplot(aes(x = cut, 
             y = price)) +
  geom_boxplot(aes(fill = "cut", 
                   color = "cut"),
               alpha = 0.4)
```

What more can you say from this visualization than the one you made in question 9?

*Once again, I can see the outliers, the percentiles, the median, and overall a clearer picture of the range of data.*


13. Use `geom_violin()` to create a violin plot for `carat` and `clarity`. Include a boxplot inside the violin. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = clarity,
             y = carat)) +
  geom_violin(aes(color = clarity)) +
  geom_boxplot((aes(color = clarity, fill = clarity)), col = 1, width = 0.2)
```

Summary:
  *The violin plot shows the distribution of diamond carat weights across different clarity grades. Lower clarity diamonds (like I1, SI2) tend to have a wider range and higher carat values, while higher clarity diamonds (like IF, VVS1) are generally smaller and less variable in size. This suggests that larger diamonds are more likely to have lower clarity.*


14. Use `geom_violin()` to create a violin plot for `carat` and `color`. Include a point for the mean, and error bars around the mean. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = color,
             y = carat)) +
  geom_violin(aes(color = color, fill = color)) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.5, color = "tomato") +
  stat_summary(fun = "mean", geom = "point")
```

15. Use `geom_jitter()` to create a jitter plot for `price` and `cut`. Include a point for the mean and error lines around the mean. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

Question:
  *How does the average price of diamonds vary by cut, and is there  overlap in price distributions between cut categories?*

Prediction:
  *If diamond price depends on cut quality, we should see tighter clusters of lower prices as cut quality gets worse. On the other hand, if cut has little impact on price, the data points would be more randomly spread out. A clear positive trend would mean that average prices drop consistently with lower cut grades.*


```{r}
diamonds_sample %>%
  ggplot(aes(x = cut,
             y = price)) +
  geom_jitter(aes(color = cut), alpha = 0.7) + 
  geom_boxplot(aes(color = cut), alpha = .5) + 
  stat_summary(fun = "mean", geom = "point", color = "black") + 
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .6, color = "black")
```

Summary:
  *The graph shows ideal cuts have the lowest mean price compared to the other cuts. In contrast, premium cuts had the highest mean. Other than that, dar, good, and very good seemed to be within the same range.*

16. Take your visualization from question 14 and add on jittered points to show the whole distribution of data on top of the violins. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = color,
             y = carat)) +
  geom_violin(aes(color = color, fill = color)) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.5, color = "tomato") +
  geom_jitter(alpha = .3) +
  stat_summary(fun = "mean", geom = "point")
```

17. Use `geom_jitter()` to create a strip chart for `z` and `clarity`. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = clarity,
             y = z)) +
  geom_jitter(aes(color = clarity),
              width = 0.2, 
              height = 0, 
              alpha = 0.5) +
  stat_summary(fun = mean, 
               geom = "crossbar", 
               width = 0.5, 
               color = "blue")
```

18. Use `geom_jitter()` to create a jitter plot for `price` and `color`. Include a point for the mean and error lines around the mean. Map the color of the points to `cut`. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = color,
             y = price)) +
  geom_jitter(aes(color = cut),
              size = 2, 
              height = 0, 
              alpha = 0.5) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar")
```

19. Use `geom_jitter()` to create a jitter plot for `price` and `color`. Include a point for the mean and error lines around the mean. Facet this graph by `cut`. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = color,
             y = price)) +
  geom_jitter(color = "blue",
              size = 2, 
              height = 0, 
              alpha = 0.5) +
  stat_summary(fun = "mean", geom = "point") +
  stat_summary(fun.data = "mean_se", geom = "errorbar") +  
  facet_grid(col = vars(cut))

```

20. Do you think that the graph created in 18 or 19 is more effective? Explain why.


*I think the graph for 18 is more effective since we can better see the relationship between color and price. 19 does not offer that much more information because we can see the difference between cut in 18 as well. Overall, 18 is more straightforward.*

21. Use `stat_summary_2d()` to make a raster plot for `clarity` (y), `cut` (x), and `price` (z). Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

Question:
  *Together, how do clarity and cut influence the price of a diamond?*

Prediction:
  *If they all interact with one another, we should see that if clarity and cut increase, so should the price, and vice versa.*

```{r}
diamonds_sample %>%
  ggplot(aes(y = clarity,
             x = cut,
             z = price)) + 
   stat_summary_2d(fun = "mean", 
                   geom = "tile", 
                   color = "black")
```
Summary:
  *Based on the graph, there does not appear to be any patterns with high price correlating to high cut and clarity. Very good cut and clarity IF is the highest value, while everything else seems mixed around the middle range.*


22. Create a scatter plot for `price` and `carat`. Add a line of best fit. Facet by `cut`. Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

```{r}
diamonds_sample %>%
  ggplot(aes(x = carat,
             y = price)) +
  geom_point(aes(color = carat), 
             size = 3,
             alpha = 0.5) +
  facet_grid(col = vars(cut)) +
  geom_smooth(method = "lm", se = FALSE, color = "black")

```



23. Use `stat_summary_hex()` or `stat_summary_2d()` to create a raster plot for `carat` (y), `table` (x), and `price` (z). Make any other aesthetic modifications that you think help improve the legibility or effectiveness of the graph.

Question:
  *Together, how do carat and table influence the price of a diamond?*

Prediction:
  *If they all interact with one another, we should see that if carat and table increase, so should the price, and vice versa.*

```{r}
diamonds_sample %>%
  ggplot(aes(x = table,
             y = carat,
             z = price)) +
  stat_summary_2d(fun = "mean", geom = "tile", color = "black")
             
```

Summary:
  *From the graph, we see that as the carat increases, the price also increases indicated by the lighter blue clusters at the top. On the other hand, table does not seem to have any influence. It is also important to note that there are a couple outliers.*


# Free Reign

Pick any graph you created above and try to make the worst, most illegible, visualization you can. Use whatever aesthetic modifications you want, but do so within reason (e.g., don't just set `size = 100` or something). Write the question number from the original visualization you picked (e.g., "question 3").

```{r}
# question 16

diamonds_sample %>%
  ggplot(aes(x = color,
             y = carat)) +
  geom_violin(color = "black", fill = "black", alpha = 0.8) +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = 0.5) +
  geom_jitter() +
  stat_summary(fun = "mean", geom = "point")
```


# Feedback

Please save the appropriate number to the variables below:

1. Approximately how long (in minutes) do you think it took you to complete this assignment?

```{r}
timeEstimate = 230
```

2. How difficult or challenging did you find this assignment?

Extremely easy -------------------------------------- Extremely challenging 
                                                     (impossible to complete)
    1                                                           10
    
```{r}
challenge = 6
```
    
    
3. How engaging or fun did you find this assignment?
  
Extremely Boring and Dry -------------------------------------- Extremely Engaging and Fun 
          1                                                                   10

```{r}
engagement = 6
```

4. If you have any other feedback about this assignment, please provide it below.




