---
title: ""
author: ""
date: ""
output: html_document
---


1. Use each of the following operators in a new expression:

"+"

```{r}
8 + 8
```

"-"

```{r}
6 - 2
```

"*"

```{r}
4 * 7 
```

"/"

```{r}
4 / 2
```

"^"

```{r}
6^2
```

"%%"

```{r}
10 %% 2
```


2. Use each of the following operators in a new expression:

"<"

```{r}
5 < 6
```

">"

```{r}
10 > 5
```

"<="

```{r}
3 <= 3
```

">="

```{r}
4 >= TRUE + TRUE
```

"=="

```{r}
5 == 9

```

"!="

```{r}
TRUE != FALSE
```

3. 
  A. Use each of the operators below in a new expression. 

"&"

```{r}
TRUE & TRUE
```

"|"

```{r}
TRUE | FALSE
```

"!"

```{r}
!FALSE

```

B. Use *each* operator (`&`, `!`, and `|`) combined with at least one operator from exercise 1 or 2 as well.

```{r}
TRUE + !FALSE

TRUE & FALSE == FALSE

TRUE | FALSE != 0

```


C. Now, use *each* operator (`&`, `!`, and `|`) combined with at least one *other* operator (logical or comparison operators).

```{r}
(TRUE + !FALSE) == 2
!(2==2)
(10 == 2) | (7 <=7) 
```


4. Define variables of each of the following types and use `typeof()` to verify their type.

A. number

```{r}
var1 <- 2
typeof(var1)
```

B. character

```{r}
var2 <- "Hi"
typeof(var2)
```


C. logical

```{r}
var3 <- TRUE
typeof(var3)
```

5. Create 3 vectors that each contain a different type of data (number, character, and logical). See which operations can be applied to each vector.

  A. number

```{r}
vector1 = c(2, 3, 4, 5)
vector1 * 2

```

  B. character

```{r}
vector2 = c("Hi", "Hola", "Bounjour")
vector2 == "Hi!"
```

  C. logical

```{r exercise 5}
vector3 = c(TRUE, FALSE, TRUE)
vector3 + TRUE

```

6.

  A. Index the 1st and 2nd element from the vector in 5A

```{r}
vector1[c(1, 2)]

```

  B. Index the 2nd and 3rd element from the vector in 5B.
  
```{r}
vector2[c(2, 3)]
```
  
  C. Save the first element from your first vector to a variable called `first_first`.
  
```{r}
first_first <- vector1[1]
first_first
```

  D. Save the last element from your last vector to one called `last_last`.
  
```{r}
last_last <- vector3[length(vector3)]
last_last

```
  

7. Create a dataframe called `my_df` that has information about your top 3 favorite animals. It should contain 3 columns: 1. One that has the words, "first", "second", and "third". 2. Another that has the animals names. 3. A third having a rating, on a scale of 1-10, of how likely you think it is you could beat that animal in a fight to the death.

```{r exercise 7}


```

8. Index the second row of your dataframe and save it to a variable called `df_second_row`. Index the name of your second favorite animal and save it to a variable called `second_fave`. Index the battle likelihood rating for your third favorite animal and save it to a variable called `third_rating`.

```{r exercise 8}

```

9. Rewrite the following code as properly formatted pipe chains.

```{r, eval = FALSE}
park(drive(start_car(find("keys")), to = "campus"))


brush(tooth_paste(running_water(tooth_brush)), teeth)


```


# Feedback 

Please save the appropriate number to the variables below:

1. Approximately how long (in minutes) do you think it took you to complete this assignment?

```{r}
timeEstimate = 60
```

2. How difficult or challenging did you find this assignment?

Extremely easy -------------------------------------- Extremely challenging 
                                                     (impossible to complete)
    1                                                           10
    
```{r}
challenge = 5
```
    
    
3. How engaging or fun did you find this assignment?
  
Extremely Boring and Dry -------------------------------------- Extremely Engaging and Fun 
          1                                                                   10

```{r}
engagement = 7
```

4. If you have any other feedback about this assignment, please provide it below.




