---
title: "ggplot intro practice problems"
author: "Amy"
date: "8/27/2025"
output: html_document
---

```{r}
library(tidyverse)
# Modify the file path in the code below to properly load the "diamonds_sample.RData" file
load("C:/Users/amyng/OneDrive/Documents/GitHub/codingNCML_Wocs/Amy/R Bootcamp/R Basics - 1/Intro to ggplot/diamonds_sample.RData") 
```


## Intro

You are going to work with a subset of data in the df `diamonds_sample`. All questions below will be in reference to `diamonds_sample`.

* `price` - price in USD
* `carat` - weight of the diamond
* `cut` - quality of the cut
* `color` - diamon color
* `clarity` - a measurement of how clear the diamond is
* `x` - length in mm
* `y` - width in mm
* `z` - depth in mm
* `depth` - total depth percentage
* `table` - width of top of diamond relative to widest point

## Aesthetic Setting

Let's first practice setting different aesthetics manually. In all the following graphs, use `geom_point()` to represent each individual observation with a point. After creating the visualization, write a brief sentence describing the relationship you see between the two variables.

1. Create a graph with `price` on the y-axis, `carat` on the x-axis, and change the color of all the points to your favorite color.

```{r}
library(ggplot2)
library(dplyr)

diamonds %>%
  ggplot(aes(x = carat, 
             y = price)) +
  geom_point(color = "purple")
```

2. Create a graph with `table` on the y-axis, `depth` on the x-axis, and change the color of all the points to chartreuse

```{r}
diamonds %>%
  ggplot(aes(x = depth, 
             y = table)) +
  geom_point(color = "chartreuse")
```

3. Create a graph with `carat` on the y-axis, `table` on the x-axis, and change the shape of all points to 8

```{r}
diamonds %>%
  ggplot(aes(x = table, 
             y = carat)) +
  geom_point(color = "blue",
             shape = 8)
```

4. Create a graph with `x` on the y-axis, `y` on the x-axis, and change the alpha to 0.25

```{r}
diamonds %>%
  ggplot(aes(x = x, 
             y = y)) +
  geom_point(color = "blue",
             alpha = 0.25)
```

5. Create a graph with `depth` on the y-axis, `z` on the x-axis, make the shape 13 and the color orchid

```{r}
diamonds %>%
  ggplot(aes(x = z, 
             y = depth)) +
  geom_point(shape = 13,
             color = "orchid")
```

6. Create a graph with `table` on the y-axis, `carat` on the x-axis, make the color tomato, shape 3, size 2, and adjust the alpha to a level you think is best

```{r}
diamonds %>%
  ggplot(aes(x = carat, 
             y = table)) +
  geom_point(color = "tomato",
             shape = 3,
             size = 2,
             alpha = 0.5)
```

(Challenge) Create a graph with `price` on the y-axis, `depth` on the x-axis, make each observation be represented by a smaller gold circle within a larger magenta circle outline (hint: think about your layering order).

```{r}
diamonds %>%
  ggplot(aes(x = depth, 
             y = price)) +
  geom_point(color = "gold",
             size = 2) +
  geom_point(color = "magenta",
             size = 2.5,
             shape = 1)
```


## Aesthetic Mapping

Now, instead of manually setting the aesthetics you will map them to variables in the dataset. In doing so, you will be able to explore these relationships even further! You will be given instructions on how to generate a particular graph.

Before running the code to create your graph...
A. Write down the *question* that this graph is exploring.
B. Write a brief *prediction* of what you expect the relationship or pattern to be. 

After observing the visualization... 
A. Write a brief *summary* describing what that relationship or pattern actually is, and explain what evidence you have for this (e.g., "all the brown colored dots that correspond to 'Ideal' cut diamonds are higher on the y-axis, which here is 'price'. So I can see that 'Ideal' cut diamonds tend to be the most expensive.")

*NOTE: For some of these visualizations, it may be hard to discern any actual patterns or relationships. Just try your best to describe what you see. If it is too hard to determine anything, state that and explain why!*

1. Create a graph with `price` on the y-axis, `carat` on the x-axis, and map the color of all the points to `cut`. Make any other aesthetic modifications you need to make the visualization ans legible as possible.

Question:

  *How are the price and carat of varying cuts related?*


Prediction:

  *If the price of a diamond rises as the carats in a cut increases, we would see a positive and linear trend for the line of best fit.*
  
  *If that is not the case that there is a positive relationship between price and carat, there would be a negative trend for the line of best fit.*

```{r}
diamonds_sample %>%
  ggplot(mapping = aes(x = carat,
                         y = price,
                         color = cut)) +
  geom_point(aes(color = cut),
             shape = 15,
             alpha = .5,
             size = 3) + 
  geom_smooth(method = "lm", se = FALSE, color = "blue")
```

Summary:

  *In summary, there is a positive linear trend between the price and carats in a cut. More specifically, the higher the carat, the higher the price. Fair cuts of diamonds seem to be priced lower than Good, Very Good, Premium, Ideal consistently.*


2. Create a graph with `carat` on the y-axis, `z` on the x-axis, and map the color of all the points to `color`. Make any other aesthetic modifications you need to make the visualization ans legible as possible.

Question:

  *How does the depth of a diamond correlate to the carats, while also considering the colors of the diamond?*


Prediction:

  *If the depth of a diamond increases, the carats will also increase, resulting in a positive linear line of best fit. If the color also has an affect, there will also be distinct clusters representative of the individual colors.*
  
  *If the depth decreases, the carats will also decrease, resulting in a negative trend. If color also does not influence carat or depth, there should be no noticeable clusters.*


```{r}
diamonds_sample %>%
  ggplot(mapping = aes(x = z,
                       y = carat,
                       color = color)) +
  geom_point(aes(color = color),
             shape = 15,
             alpha = .5,
             size = 5) + 
  geom_smooth(method = "lm", se = FALSE)
```

Summary:

  *In summary, depth and carat seem to have a positive linear line of fit and relationship across all colors of diamonds, and there does not appear to be significant distinct clusters of colors. This indicates color does not have an effect. However, there was one major outlier: A "G" diamond had 1 carat with a depth of 0mm.*

3. Create a graph with `depth` on the y-axis, `y` on the x-axis, and map the color of all the points to `clarity`. Make any other aesthetic modifications you need to make the visualization ans legible as possible.

Question:

  *How is depth correlated to width of diamonds of different clarity?*


Prediction:

  *If the depth of a diamond increases, the width will also increase, resulting in a positive linear line of best fit. If clarity of the diamond affects either depth or width, their lines of best fit should be different or there should be different colors.*
  
  *If depth and width have no relationship, they should not have any noticeable trends. If clarity of diamond has no effect, there will be no noticable clusters or differing lines of best fit.*

```{r}
diamonds_sample %>%
  ggplot(mapping = aes(x = y,
                       y = depth,
                       color = clarity)) +
  geom_point(aes(color = color),
             shape = 15, 
             alpha = 0.5, 
             size = 3) +
  geom_smooth(method = "lm", se = FALSE)
```

Summary:

  *In summary, the lines of best fit are roughly linear, which represents that width does not correlate with depth as there is no trends. The data points cluster around depths of 60 to 65 and y of 4 to 8.5. There are no noticeable clusters of diamond clarity, showing that clarity does not affect depth or width.*

## Free Reign

Below, try to make the worst, most illegible, visualization you can. Use whatever aesthetic modifications you want, but do so within reason (e.g., don't just set `size = 100` or something).

4. Create a graph with `price` on the y-axis, `carat` on the x-axis.

```{r}
diamonds_sample %>%
  ggplot(mapping = aes(y = carat,
                       x = x,
                       color = color,)) +
  geom_point(aes(color = cut),
             shape = "triangle",
             size = 4, 
             alpha = .1) + 
  geom_point(aes(color = "tomato"),
             shape = 4,
             size = 1,
             alpha = .9)
```


# Feedback

Please save the appropriate number to the variables below:

1. Approximately how long (in minutes) do you think it took you to complete this assignment?

```{r}
timeEstimate = 150
```

2. How difficult or challenging did you find this assignment?

Extremely easy -------------------------------------- Extremely challenging 
                                                     (impossible to complete)
    1                                                           10
    
```{r}
challenge = 7
```
    
    
3. How engaging or fun did you find this assignment?
  
Extremely Boring and Dry -------------------------------------- Extremely Engaging and Fun 
          1                                                                   10

```{r}
engagement = 6
```

4. If you have any other feedback about this assignment, please provide it below.




